---
title: "Loan Default Prediction Using SVM"
author: "Joseph Oladokun"
date: "October 23, 2018"
output: html_document
---
```{r}
library(ggplot2)
library(caTools)
library(e1071)
```

Structure of the Data gotten from LendingClub.com
```{r }
loans <- read.csv('loan_data.csv')
str(loans)
```

Summary of the Data

```{r loans}
summary(loans)
```
Convert the following this columns to categorical data using factor()

inq.last.6mths
delinq.2yrs
pub.rec
not.fully.paid
credit.policy

```{r }
loans$credit.policy <- factor(loans$credit.policy)
loans$inq.last.6mths <- factor(loans$inq.last.6mths)
loans$delinq.2yrs <- factor(loans$delinq.2yrs)
loans$pub.rec <- factor(loans$pub.rec)
loans$not.fully.paid <- factor(loans$not.fully.paid)
```

Exploratory Data Analysis

I will use ggplot2 to visualise the data.

I will create a histogram of fico scores colored by column not.fully.paid

```{r}
pl <- ggplot(loans,aes(x=fico)) 
pl <- pl + geom_histogram(aes(fill=not.fully.paid),color='black',bins=40,alpha=0.5)
pl + scale_fill_manual(values = c('green','red')) + theme_bw()
```

Create a barplot of purpose counts, colored by not.fully.paid. Use position=dodge in the geom_bar argument.
```{r}
pl <- ggplot(loans,aes(x=factor(purpose))) 
pl <- pl + geom_bar(aes(fill=not.fully.paid),position = "dodge")
pl + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
Create a scatterplot of fico score versus int.rate. Does the trend make sense? Play around with the color scheme if you want.
```{r}
ggplot(loans,aes(int.rate,fico)) +geom_point(aes(color=not.fully.paid),alpha=0.3) + theme_bw()
```

Building the SVM Model for the prediction

Split the data into training and test sets using the caTools
```{r}
set.seed(101)
spl = sample.split(loans$not.fully.paid, 0.7)
train = subset(loans, spl == TRUE)
test = subset(loans, spl == FALSE)

```
Now we use the svm() function to train the model using the training set
```{r}
model <- svm(not.fully.paid ~ .,data=train)
```
Get the summary of the model
```{r}
summary(model)
```
Use "predict" to predict new values from the test set using the model above
```{r}
predicted.values <- predict(model,test[1:13])
table(predicted.values,test$not.fully.paid)
```
We got some bad results! With the model classifying everything into one group! Let's tune our model to try to fix this.

Tuning the model
```{r}
tune.results <- tune(svm,train.x=not.fully.paid~., data=train,kernel='radial',
                  ranges=list(cost=c(1,10), gamma=c(0.1,1)))

model <- svm(not.fully.paid ~ .,data=train,cost=10,gamma = 0.1)
predicted.values <- predict(model,test[1:13])
table(predicted.values,test$not.fully.paid)
```

